\chapter{User Study Tasks}

\section*{Script:}

I will now hand you a task description. You can take your time to read it, I will begin timing you when you’re ready to attempt the task. For each task, you will find a corresponding Microsoft Excel data file which you must use in completing the task if required.

\section*{Task A:}

You work in a government agency and are trying to see how population is growing in 4 different countries. The ‘Task A’ file contains information regarding how many births and deaths occurred in 1 year in each country (the numbers are scaled to account for population). Use the Sketchography app to create a column bar chart that will allow someone to compare birth and death rates across the countries.

Speak out aloud the name of the country where the population is declining (where the death rate is larger than the birth rate).

\section*{Task B:}

You are comparing universities to see which are the most selective. The ‘Task B’ file contains information regarding how many students applied to each of 4 universities, and how many were admitted. Use the Sketchography app to create a column bar chart that will allow someone to compare applications and admittances across the universities.

Speak out aloud the name of the university that had the highest number of applicants.

\section*{Task C:}

In Sketchography, open the formal chart you created in Task B.

Make the bars shorter in height without changing the size of the graph itself.

\section*{Task D:}

The ‘Task D’ file contains the same data about universities as Task B, as well as an Excel chart similar to the one you created in Sketchography in Task B.

Make the bars shorter in height without changing the size of the graph itself.